% --------------------------------------------------------------------------------
\chapter{History of model changes}
% --------------------------------------------------------------------------------

The forecasting environment, which is composed of the ICON model and the data assimilation system, 
is subject to continuous improvements and modifications. The most important ones in terms of forecast 
quality and output products are depicted below. For additional information, the reader is referred to the 
official change notifications which are available from
\begin{note}
\url{http://www.dwd.de/DE/fachnutzer/forschung_lehre/numerische_wettervorhersage/nwv_aenderungen/nwv_aenderungen_node.html}
\end{note}


% Timeline:
%
% The macro "\colorA" corresponds to ICON model changes
%           "\colorB" corresponds to data assimilation changes.
%
\begin{vtimeline}[description={text width=0.85\textwidth}, 
 row sep=7ex, 
 add bottom line,
 line offset=50pt,
 timeline color=colorBlue,
 timeline color2=colorRed]
\colorA  2015-01-16 & First operational ICON forecast run at a horizontal resolution of $13\,\mathrm{km}$ and $90$ vertical levels. Model top at $75\,\mathrm{km}$\endlr
\colorA  2015-02-25 & $4$ additional forecast runs 03, 09, 15, 21\,UTC up to $30\,\mathrm{h}$ lead time. Maximum forecast lead times for 06 and 18\,UTC runs extended to $120\,\mathrm{h}$\endlr
\colorA  2015-03-04 & Improved wind gust diagnostic for mountainous regions \endlr
%        2015-06-17 & Introduction of a surface wind dependent Charnock-Parameter.
\colorA  2015-07-07 & Time interval over which max/min temperatures \texttt{TMAX\_2M}, \texttt{TMIN\_2M} are collected changed to $6\,\mathrm{h}$ (formerly $3\,\mathrm{h}$). 
                      Time interval over which maximum wind gusts \texttt{VMAX\_10M} are collected changed to $1\,\mathrm{h}$ (formerly $3\,\mathrm{h}$).\endlr  
\colorA  2015-07-21 & Launch of the ICON-EU nest with a horizontal resolution of $6.5\,\mathrm{km}$ and $50$ vertical levels. Model top at $50\,\mathrm{km}$.\endlr
\colorB  2015-09-02 & Assimilation of selected microwave satellite radiance channels over land (AMSU-A and ATMS), which have so far been assimilated only over the oceans.\endlr
\colorA  2015-12-01 & Surface tile approach for land, sea-ice and lake points, including snow-tiles. Each grid point can have up to $5$ tiles consisting of 3 land tiles 
                      (dominant land-use types), a lake tile or a sea-water plus sea-ice tile. Land tiles may have additional snow tiles.\endlr
\colorB  2016-01-20 & Launch of the ensemble data assimilation system (LETKF, Localized Ensemble Transform Kalman Filter), providing a $40$ member analysis ensemble at R2B6N7. 
                      For deterministic forecasts the 3D-Var assimilation system is replaced by En-Var (Ensemble Variational analysis system) which makes use 
                      of the ensemble-based model error covariances.\endlr
\end{vtimeline}